---
title: "pubh7462_hw2_RMD"
author: Tsz Fung Wong
date: January 28,2022
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
#Load the good stuff
library(tidyverse)
library(DataExplorer)
library(ggplot2)
library(gt)
#Working directory for .RMD
knitr::opts_knit$set(echo = TRUE,
root.dir = rprojroot::find_rstudio_root_file())
#Controlling figure output in markdown
knitr::opts_chunk$set(
# fig.height = 4
fig.width = 6,
# fig.asp = .5,
out.width = "90%",
# out.height =
fig.align = "center",
cache = FALSE,
echo = TRUE
)
#Set Theme for ggplot2
theme_set(theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom"))
#Set Scientific notation output for knitr
options(scipen = 999)
options(digits = 4)
```

# Question 1
```{r message = FALSE, warning = FALSE}
brfss = read_csv("./data/brfss_smart_2010.csv")
brfss_df = as_tibble(brfss) %>%
  janitor::clean_names() %>% #clean names
  filter(topic == "Overall Health") %>%
  select(c("year", "locationabbr", "locationdesc", "response", 
             "sample_size", "response", "data_value")) %>%
    mutate(county = stringr::str_remove(locationdesc, str_flatten(str_c(unique(locationabbr), " - "), "|"))) %>%
  select(-"locationdesc") %>%
  rename("state" = "locationabbr", 
           "prop_of_response" = "data_value") 
brfss_df$response = fct_relevel(factor(brfss_df$response), "Excellent", "Very good", "Good", "Fair")
brfss_df$county = factor(brfss_df$county)
brfss_df$state = factor(brfss_df$state)

```
# Data Description
The modified BRFSS data consist of `r nrow(brfss_df)` observations described with `r ncol(brfss_df)` variables measuring overall health conditions. The first variable year indicate the time of data collection, ranging from 2002 to 2010. Variable state and county indicate the geographical information of the observation from `r n_distinct(brfss_df$state)` states. Response variable indicate the `r n_distinct(brfss_df$response)` type of response indicating overall health conditions. Sample_size variable indicate the amount of response collected from the specific county for different type of response. Prop_of response indicate the proportion of response accociated with the corresponding type of response. 


# Q3.3.1 
In year 2004, the states that were observed at 6 locations are `r names(which(table(str_sub(unique(brfss$Locationdesc), 1, 2)) == 6))`. 

# Q3.3.2
```{r message = FALSE, warning = FALSE}
brfss_plot = brfss_df %>%
  filter(year >= '2002' & year <= '2010') %>%
  group_by(year, state) %>%
  summarise(num_location = n_distinct(county)) %>%
  mutate(
    year = year %>%
           as.character() %>%
           as.numeric()
  ) 

brfss_plot %>%
  ggplot(aes(x = year, y = num_location, colour = fct_reorder2(state, year, num_location))) +
  geom_line(size = 1, alpha = 0.44) +
  stat_smooth(alpha = 0.1, 
              size = 0.2, 
              method = "loess",
              span = 2,
              se = FALSE) +
  labs(
    x = "Year",
    y = "Number of Observed Locations",
    title = "Number of Observed Locations by State from 2002 - 2010"
  ) +
  scale_colour_viridis_d("State") +
  theme(legend.position = "right", 
        legend.key.size = unit(0.2, "cm"),
        legend.key.height = unit(0.2, 'cm'),
         legend.key.width = unit(0.05, 'cm'))

```
State NJ has the highest mean number of locations. Most state has low number of locations observed over time, whereas state FL has two peaks only in year 2007 and 2010.

# Q3.3.3
```{r message = FALSE, warning = FALSE}
brfss_sum = brfss_df %>%
  filter(year %in% c("2002", "2010", "2006")) %>%
  filter(state %in% c("MN")) %>%
  filter(response %in% c("Excellent", "Good", "Poor")) %>% 
  group_by(year, response) %>%
  summarise(
    across(
      contains(c("sample", "prop")), 
      list(mean = mean, sd = sd), na.rm=TRUE, 
      .names = "{.col}_{.fn}"
    )
  )

brfss_sum  %>%
  gt() %>%
  tab_header("Summary of Sample Size and Proportion of Response")

```
More above average overall health people are sampled, and they tends to have higher response rate than the people with poor health condition. The number of people sampled increase by year generally but the proportion of people respond tends to be consistent throughout the three year.  

# Q3.3.4
Response Trend of MN
```{r message = FALSE, warning = FALSE} 
brfss_sum %>% 
  pivot_longer(-c(year, response),
  names_to = "type",
  values_to = "value"
  ) %>%
  ggplot(aes(x = year, y = value, colour = response)) +
  geom_point(size = 4, alpha = 0.44, shape = 16) +
  stat_smooth(alpha = 0.1, 
              size = 1.2, 
              method = "loess",
              span = 2,
              se = FALSE) +
  labs(
    x = "Year",
    y = "Density",
    title = "Trend of Response Portion and Sample Size by Response Type"
  ) +
  scale_colour_viridis_d("Response") +
  facet_wrap(~ type, scales = "free", ncol = 2) + 
  scale_x_continuous(breaks = c(2002, 2006, 2010))
  
```

